---
title: "p8105_hw3_zc2610"
author: "Zuoqiao Cui"
date: "2022-10-07"
output: github_document
---
```{r}
library(tidyverse)
```
## Problem 2
Import data
```{r}
acc_df = read.csv("./data/accel_data.csv") %>% 
  janitor::clean_names()
```
Combine all activity columns
```{r}
acc_df = acc_df %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_minute",
    values_to = "counts",
    names_prefix = "activity_"
  )
```
Create a weekday vs weekend variable
```{r}
acc_df = acc_df %>% 
  mutate(
    weekday_or_weekend = case_when(
      day == "Monday" ~ "weekday",
      day == "Tuesday" ~ "weekday",
      day == "Wednesday" ~ "weekday",
      day == "Thursday" ~ "weekday",
      day == "Friday" ~ "weekday",
      day == "Saturday" ~ "weekend",
      day == "Sunday" ~ "weekend",
      TRUE     ~ ""
    )
  ) 
```

```{r}
colnames(acc_df)
nrow(acc_df)
ncol(acc_df)
```

Description

1. The resulting dataset contains following variables: `r colnames(acc_df)`

2. There are totally `r nrow(acc_df)` observations and `r ncol(acc_df)` variables in the resulting dataset

Create a variable that sums up counts of activities in each day and show in a table
```{r}
acc_df %>% 
  group_by(week,day_id,day) %>% 
  summarise(
    total_activity = sum(`counts`)
  ) %>% 
  knitr::kable(digits = 1)
```

Single panel plot that shows the 24-hour activity time courses for each day

```{r}
acc_df %>% 
   ggplot(aes(x = activity_minute,y = counts,color = day)) +
   geom_point() +
   labs(
    title = "24-Hour Activity Time Courses For Each Day",
    x = "Time(minute)",
    y = "Activity Counts"
  )
```

## Problem 3
Load data
```{r}
library(p8105.datasets)
data("ny_noaa")
```
Clean data
1. Separate date into year, month, day
```{r}
ny_noaa = ny_noaa %>% 
  janitor::clean_names() %>% 
    separate(date,into = c("year","month","day"),sep = "-")
ny_noaa
```
2. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units.
Since temperature's unit is tenths of degrees and precipitation's unit is tenths of mm, I change the temperature's unit to degrees and the precipitation's unit to mm

```{r}
ny_noaa = ny_noaa %>% 
  mutate(
    prcp = prcp/10,
    tmax = as.double(tmax)/10,
    tmin = as.double(tmin)/10
  )
```
3. Find the most commonly observed values for snowfall
```{r}
total_num_of_snowfall = (nrow(ny_noaa))
num_of_na_snowfall = sum(is.na(ny_noaa$snow))
percentage_of_na = num_of_na_snowfall/total_num_of_snowfall
num_of_zero_snowfall = nrow(filter(ny_noaa,snow == 0))
percentage_of_zero = num_of_zero_snowfall/total_num_of_snowfall
```
Answer:

The percentage of NA value in snowfall column is `r sum(is.na(ny_noaa$snow))/(nrow(ny_noaa))` and the percentage of 0 value in snowfall column is `r nrow(filter(ny_noaa,snow == 0))/(nrow(ny_noaa))`.

Since the percentage of 0 value in snowfall column is `r nrow(filter(ny_noaa,snow == 0))/(nrow(ny_noaa))` > 0.5, which means more than half of the value for snowfall is 0. Therefore, the most commonly observed values for snowfall is 0.

Change month numbera to month name
```{r}
ny_noaa = ny_noaa %>% 
  mutate(
     month = recode(month, "01" = "January",
                  "02" = "February",
                  "03" = "March",
                  "04" = "April",
                  "05" = "May",
                  "06" = "June",
                  "07" = "July",
                  "08" = "Augest",
                  "09" = "September",
                  "10" = "October",
                  "11" = "November", 
                  "12" = "December"
                  )
     )
```







